Bricking up backdoors in Intrusion Detection Systems

### 2019-09-16 paper/paper.tex
Added references for scikit-learn and pytorch

\paragraph{\gls{ml} Techniques}
%\subsubsection{Deep Learning}
We used PyTorch to build an \gls{mlp}. 
% neural network classifier.
For this we used 5 fully connected layers with 512 neurons each. We applied \gls{relu} activation after each layer and dropout with a probability of 0.2 and used binary cross entropy  as loss function for training. % for the classification task itself.

%\subsubsection{\glspl{rf}}
In addition to the \gls{mlp}, we trained an \gls{rf} classifier to detect attacks in network traffic. We used the RandomForestClassifier implementation from scikit-learn for this task and used 100 estimators for training the \gls{rf}.

@@

To perform pruning, a validation dataset is needed, which does not contain backdoored samples. We take a validation dataset that is $\frac{1}{4}$ of the training dataset. We use the validation dataset for pruning as described in the next sections and a test dataset that is also $\frac{1}{4}$ to verify whether the backdoor can successfully be removed and how much the accuracy on the original data suffers. Training, validation and test dataset are pairwise disjoint.

@@

% AH. moved 'by their average activation' to beginning of sentence to make clear it applies to both variants
We implemented two variants of the pruning defense \cite{gu_badnets:_2017}: Pruning neurons by their average activation in \begin{enumerate*}\item all layers and \item only in the last layer \end{enumerate*}. Neurons with the smallest average activation (the least used ones) are pruned first. For this purpose, we look at the value of the activation function
%(\gls{relu} in our case)
in the corresponding layer for all samples in the validation dataset and prune neurons by setting the weight and bias to 0 in the layer preceding the activation function.

@@

%We propose a defense approach that can be used for similar use cases like the mechanisms recently proposed to remove backdoors from \glspl{cnn}, discussed above. 
We developed a defense approach specifically for \gls{rf} classifiers.
The concept behind our pruning defense is that leaves that are never used by samples in the validation dataset might be used by backdoored inputs. If these ``useless'' leaves are removed, performance of the classifier on the validation dataset should not decrease
%by a lot
significantly,
while decisions that are used for the backdoor are likely to be removed.
We developed several variants of our pruning defense with increasing levels of sophistication:
\begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
\item Pruning leaves based on their usage: This means that the least used leaf is pruned first and the most used one last.
\item Like (1) but consider only ``harmless'' leaves: We assume that attackers want malicious samples appear harmless.
\item Like (2) but additionally use depth to decide when
%two % AH could be more than two
leaves are used equally often: The rationale is that we assume that hand-crafted backdoors require fewer rules to classify than regular samples and thus have lower depth.
\end{enumerate}

\autoref{fig:pruningExample} shows an example of variant (3) of pruning applied to a tree
\begin{figure}[h]
\includegraphics[width=\columnwidth]{../prune_CAIA_backdoor_15/prune_oh_d.pdf}
\caption{Pruning an \gls{rf} for \unsw{}.}
\label{fig:performancePruningRF}
\end{figure}
%\todo{Max: Which pruning algorithm actually?}
and \autoref{fig:performancePruningRF} shows the resulting accuracy for pruning according to variant (3) for the \unsw{} dataset. With sufficient leaves being pruned, the accuracy of the backdoor approaches zero. The accuracy for the regular data does not decrease, it might even increase by
%thanks to the the pruning also
reducing overfitting. %With a significantly smaller validation set, pruning takes longer to remove the backdoor
%until it works
%but it still works. This is an important finding as a large validation dataset makes it hard for users without access to expensive hardware to perform the pruning.
Even with a very small validation dataset of just 1\% of its original size, the pruning still works as expected.

@@

For \gls{rf} classifiers, which can be considered one of the most important classifiers for \glspl{ids}, the pruning technique we proposed is able to reduce backdoor efficacy significantly. At the same time, the classifier's detection performance is not substantially reduced. Thus, we recommend always including a validation dataset when providing a \gls{dt} or \gls{rf} to another party. Even if the validation set is significantly smaller than the training set, the defensive properties are still upheld.

>>>>

\paragraph{\gls{ml} Techniques}
%\subsubsection{Deep Learning}
We used PyTorch~\cite{paszke2017automatic} to build an \gls{mlp}. 
% neural network classifier.
For this we used 5 fully connected layers with 512 neurons each. We applied \gls{relu} activation after each layer and dropout with a probability of 0.2 and used binary cross entropy  as loss function for training. % for the classification task itself.

%\subsubsection{\glspl{rf}}
In addition to the \gls{mlp}, we trained an \gls{rf} classifier to detect attacks in network traffic. We used the RandomForestClassifier implementation from scikit-learn~\cite{pedregosa_scikit-learn:_2011} for this task and used 100 estimators for training the \gls{rf}.

@@

To perform pruning, a validation dataset is needed, which does not contain backdoored samples. We take a validation set that is $\frac{1}{4}$ of the training set. We use the validation set for pruning as described in the next sections and a test set that is also $\frac{1}{4}$ to verify whether the backdoor can successfully be removed and how much the accuracy on the original data suffers. Training, validation and test sets are pairwise disjoint.

@@

% AH. moved 'by their average activation' to beginning of sentence to make clear it applies to both variants
We implemented two variants of the pruning defense \cite{gu_badnets:_2017}: Pruning neurons by their average activation in \begin{enumerate*}\item all layers and \item only in the last layer \end{enumerate*}. Neurons with the smallest average activation (the least used ones) are pruned first. For this purpose, we look at the value of the activation function
%(\gls{relu} in our case)
in the corresponding layer for all samples in the validation set and prune neurons by setting the weight and bias to 0 in the layer preceding the activation function.

@@

%We propose a defense approach that can be used for similar use cases like the mechanisms recently proposed to remove backdoors from \glspl{cnn}, discussed above. 
We developed a defense approach specifically for \gls{rf} classifiers.
The concept behind our pruning defense is that leaves that are never used by samples in the validation set might be used by backdoored inputs. If these ``useless'' leaves are removed, performance of the classifier on the validation set should not decrease
%by a lot
significantly,
while decisions that are used for the backdoor are likely to be removed.
We developed several variants of our pruning defense with increasing levels of sophistication:
\begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
\item Pruning leaves based on their usage: This means that the least used leaf is pruned first and the most used one last.
\item Like (1) but consider only ``harmless'' leaves: We assume that attackers want malicious samples appear harmless.
\item Like (2) but additionally use depth to decide when
%two % AH could be more than two
leaves are used equally often: The rationale is that we assume that hand-crafted backdoors require fewer rules to classify than regular samples and thus have lower depth.
\end{enumerate}

\autoref{fig:pruningExample} shows an example of variant (3) of pruning applied to a tree
\begin{figure}[h]
\includegraphics[width=\columnwidth]{../prune_CAIA_backdoor_15/prune_oh_d.pdf}
\caption{Pruning an \gls{rf} for \unsw{}.}
\label{fig:performancePruningRF}
\end{figure}
%\todo{Max: Which pruning algorithm actually?}
and \autoref{fig:performancePruningRF} shows the resulting accuracy for pruning according to variant (3) for the \unsw{} dataset. With sufficient leaves being pruned, the accuracy of the backdoor approaches zero. The accuracy for the regular data does not decrease, it might even increase by
%thanks to the the pruning also
reducing overfitting. %With a significantly smaller validation set, pruning takes longer to remove the backdoor
%until it works
%but it still works. This is an important finding as a large validation dataset makes it hard for users without access to expensive hardware to perform the pruning.
Even with a very small validation set of just 1\% of its original size, the pruning still works as expected.

@@

For \gls{rf} classifiers, which can be considered one of the most important classifiers for \glspl{ids}, the pruning technique we proposed is able to reduce backdoor efficacy significantly. At the same time, the classifier's detection performance is not substantially reduced. Thus, we recommend always including a validation set when providing a \gls{dt} or \gls{rf} to another party. Even if the validation set is significantly smaller than the training set, the defensive properties are still upheld.
